---
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: yes
    toc_float: no
---


<br>
<hr>
<br>


```{r}
# SET TABLE PARAMETERS FOR THE REPORT;
# ALSO SEE DATA STEP BELOW

table.name <- "SL-P01-T01-EXCESS-BENEFIT-TRANSAC"


### Use two levels of the path 
### to ensure the table header
### is unique (headers can be
### reused in schedules).
###
### Use a double-slash at the
### start of the table header: 
###
### "//Form990PartVIII/ProgramServiceRevenue"

#table.headers <- NULL

 table.headers <- 
  c("//IRS990ScheduleL/DQPTable",
    "//IRS990ScheduleL/DisqualifiedPersonExBnftTrGrp",
    "//IRS990ScheduleL/Form990ScheduleLPartI/DQPTable")

folder.name <- paste0( "TABLE-", table.name )
sample.size <- 200
```

---
title: "TABLE NAME: `r table.name`"
---






# Setup


```{r setup, include=FALSE}
knitr::opts_chunk$set( echo=TRUE, message=F, warning=F, fig.width=10 )
```



## Load Packages

```{r, echo=F, eval=F}
# RUN ONCE TO INSTALL PACKAGES 

devtools::install_github( 'ultinomics/xmltools' )
devtools::install_github( 'nonprofit-open-data-collective/irs990efile' )

package.list <- 
  c( "purrr","xml2","XML",
     "dplyr","pander",
     "httr","data.tree",
     "networkD3","igraph")

install.packages( package.list  )
```


```{r}
library( irs990efile )  # functions for parsing efile XMLs
library( dplyr )        # data wrangling
library( purrr )        # data wrangling 
library( pander )       # formatting
library( knitr )        # formatting
library( DT )           # table printing
library( httr )         # web requests 
library( xmltools )     # xml utilities
library( xml2 )         # xml utilities
library( XML )          # xml utilities 
library( data.tree )    # network visualization
library( networkD3 )    # network visualization
library( igraph )       # network visualization

source( "R/rdb-functions-v2.R")
source( "R/utils.R")
```





## Data

```{r}
#######
#######  DATA USED IN THE REPORT:
#######  UPDATE TO MATCH TABLE
#######




# IF THE TABLE IS IN THE 990/990EZ MAIN FORM

#index <- 
#  tinyindex %>% 
#  dplyr::filter( FormType %in% c("990","990EZ") )




# IF THE TABLE IS IN A SCHEDULE: 
 index <- read.csv( "schedule-samples/SCHEDL.csv" )

###  ALL SCHEDULES
###
#   "SCHEDA", "SCHEDB", "SCHEDC", 
#   "SCHEDD", "SCHEDE", "SCHEDF", 
#   "SCHEDG", "SCHEDH", "SCHEDI", 
#   "SCHEDJ", "SCHEDK", "SCHEDL", 
#   "SCHEDM", "SCHEDN", "SCHEDO", 
#   "SCHEDR"




head( index )
```



## The Concordance File

The **concordance** is an RDB to XML Crosswalk that maps all xpaths onto a specified set of variables. It is located in the **irs990efile** package. 

<br> 

The full data dictionary is available on GitHub. 

[**DATA DICTIONARY**](https://nonprofit-open-data-collective.github.io/irs990efile/data-dictionary/data-dictionary.html)

<br>

```{r, echo=F}
column.order <- 
  c("variable_name", "xpath", "description", "location_code_xsd", 
   "location_code_family", "location_code", "form", "form_type", 
   "form_part", "form_line_number", "variable_scope", "data_type_xsd", 
   "data_type_simple", "rdb_relationship", "rdb_table", "required", 
   "versions", "latest_version", "duplicated", "current_version", 
   "production_rule", "validated")

concordance <- 
  concordance[ , column.order ]  %>% 
  arrange( variable_name )
```


```{r}
concordance %>% 
  filter( rdb_table == table.name )  
```





```{r eval=F, echo=F}
rdb.tables <- concordance$rdb_table
unique( rdb.tables[ concordance$rdb_relationship == "MANY" ] )

#  [1] "F9-P03-T01-PROGRAMS"                           
#  [2] "F9-P07-T02-CONTRACTORS"                        
#  [3] "F9-P07-T01-COMPENSATION"                       
#  [4] "F9-P07-T01-COMPENSATION-HCE-EZ"                
#  [5] "F9-P99-T01-COMPENSATION"                       
#  [6] "F9-P08-T01-REVENUE-PROGRAMS"                   
#  [7] "F9-P08-T02-REVENUE-MISC"                       
#  [8] "F9-P09-T01-EXPENSES-OTHER"                     
#  [9] "SA-P01-T01-PUBLIC-CHARITY-STATUS"              
# [10] "SA-P99-T01-PUBLIC-CHARITY-STATUS"              
# [11] "SC-P01-T01-POLITICAL-ORGS-INFO"                
# [12] "SD-P07-T01-INVESTMENTS-OTH-SECURITIES"         
# [13] "SD-P08-T01-INVESTMENTS-PROG-RLTD"              
# [14] "SD-P09-T01-OTH-ASSETS"                         
# [15] "SD-P10-T01-OTH-LIABILITIES"                    
# [16] "SF-P01-T01-FRGN-ACTS-BY-REGION"                
# [17] "SF-P02-T01-FRGN-ORG-GRANTS"                    
# [18] "SF-P03-T01-FRGN-INDIV-GRANTS"                  
# [19] "SG-P01-T01-FUNDRAISERS-INFO"                   
# [20] "SG-P02-T01-FUNDRAISING-EVENTS"                 
# [21] "SH-P04-T01-COMPANY-JOINT-VENTURES"             
# [22] "SH-P05-T01-HOSPITAL-FACILITY"                  
# [23] "SH-P05-T02-NON-HOSPITAL-FACILITY"              
# [24] "SI-P02-T01-GRANTS-US-ORGS-GOVTS"               
# [25] "SI-P03-T01-GRANTS-US-INDIV"                    
# [26] "SJ-P02-T01-COMPENSATION-DTK"                   
# [27] "SK-P01-T01-BOND-ISSUES"                        
# [28] "SK-P02-T01-BOND-PROCEEDS"                      
# [29] "SK-P03-T01-BOND-PRIVATE-BIZ-USE"               
# [30] "SK-P04-T01-BOND-ARBITRAGE"                     
# [31] "SK-P05-T01-PROCEDURE-CORRECTIVE-ACT"           
# [32] "SK-P99-T01-BOND-PRIVATE_BIZ_USE"               
# [33] "SL-P01-T01-EXCESS-BENEFIT-TRANSAC"             
# [34] "SL-P02-T01-LOANS-INTERESTED-PERS"              
# [35] "SL-P03-T01-GRANTS-INTERESTED-PERS"             
# [36] "SL-P04-T01-BIZ-TRANSAC-INTERESTED-PERS"        
# [37] "SM-P01-T01-NONCASH-CONTRIBUTIONS"              
# [38] "SN-P01-T01-LIQUIDATION-TERMINATION-DISSOLUTION"
# [39] "SN-P02-T01-DISPOSITION-OF-ASSETS"              
# [40] "SO-T99-SUPPLEMENTAL-INFO"                      
# [41] "SR-P01-T01-ID-DISREGARDED-ENTITIES"            
# [42] "SR-P02-T01-ID-RLTD-TAX-EXEMPED-ORGS"           
# [43] "SR-P03-T01-ID-RLTD-ORGS-TAXABLE-PARTNERSHIP"   
# [44] "SR-P04-T01-ID-RLTD-ORGS-TAXABLE-CORPORATION"   
# [45] "SR-P05-T01-TRANSACTIONS-RLTD-ORGS"             
# [46] "SR-P06-T01-UNRLTD-ORGS-TAXABLE-PARTNERSHIP"  
```




<br>
<br>
<br>

<hr>

<br>
<br>
<br>


```{r, echo=F}
t.xpaths <- 
  concordance %>%
  filter( rdb_table == table.name ) %>% 
  select( xpath ) %>% 
  unique() %>% 
  arrange( xpath ) %>% 
  mutate( id=row_number() ) 

t.xpaths %>% 
  kable( align="l" )
```


```{r}
xpaths <- as.character( t.xpaths$xpath )
el <- create_edgelist_v1( xpaths )
nd <- FromDataFrameNetwork( network=el )
print( nd )
```


```{r, fig.height=20}
SetGraphStyle( nd, rankdir = "LR")
SetEdgeStyle( nd, arrowhead = "vee", color = "grey20", penwidth = 2 )
SetNodeStyle( nd, 
              style = "filled,rounded", 
              shape = "box", 
              fillcolor = "LightBlue", 
              fontname = "helvetica", 
              fontcolor="black",
              tooltip = GetDefaultTooltip )

# SetNodeStyle(acme$IT, fillcolor = "LightBlue", penwidth = "5px")
plot( nd )
```


```{r}
# old version
find_group_names( table.name )

# new version
find_table_headers( table.name )

# manual
if( ! is.null(table.headers) )
{table.headers}
```


<br>
<br>
<br>

<hr>

<br>
<br>
<br>



# Build the Table


```{r, results="hide"}
# create new folder
dir.create( folder.name )
setwd( folder.name )

# create sample
set.seed( 1234 )
sample.index <- sample_n( index, sample.size )
write.csv( index, "sample-index.csv", row.names=F )
sample.urls <- sample.index$URL

# erase existing log files
file.create("XPATH-LOG.txt") 
file.create("FAIL-LOG.txt")


start.build.time <- Sys.time()    # --------------------

results.list <- list()
xpath.list <- list()


for( i in 1:length( sample.urls ) )
{
  
  url <- sample.urls[i]
  
  results.list[[i]] <- 
    build_rdb_table_v2( url, 
                        table.name, 
                        table.headers=table.headers )
  
  xpath.list[[i]]   <- 
    get_table_xpaths( url, 
                      table.name, 
                      table.headers=table.headers )
  
  if( i %% 100 == 0 ){ print(i) }
  
}


end.build.time <- Sys.time()      # --------------------


df <- dplyr::bind_rows( results.list )


fail.log <- readLines( "FAIL-LOG.txt" )
all.xpaths <- xpath.list %>% unlist() %>% unique() %>% sort()
cat( all.xpaths, sep="\n", file="XPATH-LOG.txt" )
```

TABLE:   **`r table.name`**  

```{r, echo=F, results="hold", eval=T}
# print( paste0(  "TABLE:   ", table.name ) )
print( paste0( "NUMBER OF RETURNS PROCESSED:   ", length(sample.urls) ) )
print( paste0( "TOTAL NUMBER OF TABLE ROWS:   ", nrow(df) ) )

# NUMBER OF RETURNS PROCESSED:   **`r length(sample.urls)`**  
# TOTAL NUMBER OF TABLE ROWS:   **`r nrow(df)`**  
```

TOTAL RUN TIME: 

```{r, results='hold', echo=F}
print( end.build.time - start.build.time )
print( difftime( start.build.time, end.build.time, units="hours") )
```


## Write to File

```{r}
write.csv( df, paste0( folder.name, "/", table.name, ".csv" ), row.names=F )
head(df) 
```

Are all names standardized? 

```{r}
names(df)
```




<br>
<br>
<br>

<hr>

<br>
<br>
<br>







# Diagnostics 

## Missing or Extra XPATHS 

Check for xpaths that were used in the tables but do not exist in the Concordance file.  

```{r}
c.table <- dplyr::filter( concordance, rdb_table == table.name ) 
xpath.concordance <- c.table$xpath
setdiff( all.xpaths, xpath.concordance )
```

Capture to update the Concordance file. 

```{r}
these <- setdiff( all.xpaths, xpath.concordance )
cat( these, sep="\n", file=paste0( folder.name, "/MISSING-XPATHS.txt" ) )
```



## Failes URLs 

Check the status of failed URLs. 

```{r}
if( length(fail.log) > 0 )
{ 
  sapply( fail.log, get_message ) %>% 
  table() %>% 
  pander() 
} else {
  print("ALL URLs ARE VALID")
}
```






<br>
<br>
<br>

<hr>

<br>
<br>
<br>






# XML Table Structure


```{r, eval=F, echo=F}

### Get nodesets from test data: 

if( is.null(table.headers) )
{ table.headers <- find_table_headers( table.name ) }

all.headers <- paste0( table.headers, collapse="|" )

node.list <- list()

for( i in 1:length( test.urls ) )
{
  url <- test.urls[i]
  doc <- NULL
  try( doc <- xml2::read_xml( file(url) ), silent=T )
  if( is.null(doc) ) next
  xml2::xml_ns_strip( doc )
  node.list[[i]] <- xml2::xml_find_all( doc, all.headers, flatten=TRUE )
  # node.list[[i]] <- XML::getNodeSet( doc, all.groups )
}

### Getting xpaths from nodelists: 

get_unique_xpaths <- function(x)
{
  xpz <- 
    x %>% 
    xml_parent() %>% 
    xmltools::xml_get_paths() %>%
    unlist() %>% 
    unique()
  return(xpz)
}

all.xpaths <- 
  lapply( node.list, get_unique_xpaths ) %>% 
  unlist() %>% 
  unique()
```


## Create Network Edgelists


The **create_edgelist()** functions v1 and v2 are in R/utils.R. 

V1 retains the root nodes Return and ReturnData. 

V2 drops the root nodes and uses form versions "990" and "990EZ" as the root to visualize the sets of tables separately. 

```{r, eval=F, echo=F}
all.xpaths <- 
  readLines( paste0( folder.name,"/XPATH-LOG.txt" ) ) %>% 
  unique() %>% 
  sort()
```

```{r}
# KEEP ROOT NOTES Return & ReturnData
el <- create_edgelist_v1( all.xpaths )
```

<br>
<hr>
<br>

Edgelist format: 

```{r, echo=F}
el.demo <- 
structure(list(V1 = c("Return", "ReturnData", "IRS990", "Form990PartVIISectionAGrp", 
"Form990PartVIISectionAGrp", "Form990PartVIISectionAGrp", "Form990PartVIISectionAGrp", 
"Form990PartVIISectionAGrp", "Form990PartVIISectionAGrp", "Form990PartVIISectionAGrp"
), V2 = c("ReturnData", "IRS990", "Form990PartVIISectionAGrp", 
"PersonNm", "TitleTxt", "AverageHoursPerWeekRt", "IndividualTrusteeOrDirectorInd", 
"OfficerInd", "ReportableCompFromOrgAmt", "ReportableCompFromRltdOrgAmt"
)), row.names = c(NA, 10L), class = "data.frame") 
  
el.demo %>% pander()
```



<br>
<br>
<br>

<hr>

<br>
<br>
<br>



## Visualize Tree Structure


```{r}
# library( data.tree )
nd <- FromDataFrameNetwork( network=el )
print( nd )
```


```{r, fig.height=20}
SetGraphStyle( nd, rankdir = "LR")
SetEdgeStyle( nd, arrowhead = "vee", color = "grey20", penwidth = 2 )
SetNodeStyle( nd, 
              style = "filled,rounded", 
              shape = "box", 
              fillcolor = "LightBlue", 
              fontname = "helvetica", 
              fontcolor="black",
              tooltip = GetDefaultTooltip )

# SetNodeStyle(acme$IT, fillcolor = "LightBlue", penwidth = "5px")
plot( nd )
```




```{r}
# library( networkD3 )
# net <- ToDataFrameNetwork( nd, "name" )
# simpleNetwork( net, fontSize = 12 )
simpleNetwork( el, fontSize = 12 )
```


```{r, fig.height=10}
#plot with networkD3
radnet <- ToListExplicit( nd, unname = TRUE )
radialNetwork( radnet, fontSize = 12 )
```


```{r, fig.height=15, eval=F, echo=F}
diagonalNetwork( List = radnet, 
                 fontSize = 12, 
                 opacity = 0.9)
```





```{r,eval=F, echo=F}
## Compare 990 to 990EZ Trees

# DROP TOP TWO NODES: Return & ReturnData
el <- create_edgelist_v2( all.xpaths )

# library( igraph )
g <- graph_from_data_frame( d=el, directed=TRUE, vertices=NULL )
plot( g, layout=layout_as_tree )
```






<br>
<br>
<br>

<hr>

<br>
<br>
<br>





# XPATH Summary

## Unique Xpaths for this RDB Table

```{r}
t.xpaths <- 
  concordance %>%
  filter( rdb_table == table.name ) %>% 
  select( xpath ) %>% 
  unique() %>% 
  arrange( xpath ) %>% 
  mutate( id=row_number() ) 

t.xpaths %>% 
  kable( align="l" )
```


```{r, echo=F, eval=F}
## Compare xpath Levels

xpaths <- as.character( t.xpaths$xpath )
nodes <- strsplit( xpaths, "/" )
xpath.levels <- sapply( nodes, length )
max.lev <- max( xpath.levels )
nodes <- lapply( nodes, function(x){ c(x, rep("",max.lev-length(x) ) ) } )

# compare xpath levels
xpath.levels <- data.frame( do.call( cbind, nodes ), stringsAsFactors=F )
xpath.levels[ , 1:10 ] 
```





## Grouping Variables

Identify all unique groups used in xpaths. 

```{r}
if( is.null(table.headers) )
{ find_table_headers( table.name ) } else
table.headers
```



Groups can then be used to extract the 1:M RDB tables: 

```{r, paged.print=FALSE}
if( is.null(table.headers) )
{ table.headers <- find_table_headers( table.name ) }
demo.url <- "https://nccs-efile.s3.us-east-1.amazonaws.com/xml/201913079349300206_public.xml"
doc <- xml2::read_xml( demo.url )
xml2::xml_ns_strip( doc )
df.demo <- get_table_v2( doc, table.name, table.headers )
df.demo 
```





## Standardize Variable Names

Mapping of xml variable names to concordance variable names: 

```{r}
get_var_map( table.name=table.name )
```

A table with proper names: 

```{r paged.print=FALSE}
v.map <- get_var_map( table.name=table.name )

if( ! is.null(df.demo) )
{re_name( df.demo, v.map ) %>% as.data.frame()}
```


<br>
<br>
<br>

<hr>

<br>
<br>
<br>




## Print Table Structure

Demo of functions - demo data only. 

```{r, eval=F, echo=F}
node.list[[1]] %>% xml_parent() %>% xml2::xml_structure()
```

```
[[1]]
<Form990PartVIISectionAGrp>
  <PersonNm>
    {text}
  <TitleTxt>
    {text}
  <AverageHoursPerWeekRt>
    {text}
  <IndividualTrusteeOrDirectorInd>
    {text}
  <OfficerInd>
    {text}
  <ReportableCompFromOrgAmt>
    {text}
  <ReportableCompFromRltdOrgAmt>
    {text}
  <OtherCompensationAmt>
    {text}
```

```{r, eval=F}
node.list[[1]] %>% xml_parent() %>% xmltools::xml_view_trees()
```


```
 (1) ------------

 +-- Form990PartVIISectionAGrp
  +-- PersonNm
  +-- TitleTxt
  +-- AverageHoursPerWeekRt
  +-- IndividualTrusteeOrDirectorInd
  +-- OfficerInd
  +-- ReportableCompFromOrgAmt
  +-- ReportableCompFromRltdOrgAmt
  ++-- OtherCompensationAmt


 (2) ------------

 +-- Form990PartVIISectionAGrp
  +-- PersonNm
  +-- TitleTxt
  +-- AverageHoursPerWeekRt
  +-- IndividualTrusteeOrDirectorInd
  +-- OfficerInd
  +-- ReportableCompFromOrgAmt
  +-- ReportableCompFromRltdOrgAmt
  ++-- OtherCompensationAmt
```



















```{css, echo=F}
h1, h2{
  margin-top:80px;
  margin-bottom:20px;
}


h1{
  color: #762e0c;
  font-size: 38px;
  text-transform: uppercase;
}

h2, h3{
  color: #003169;
  font-size: 30px;
}



table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:800px;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px;
    margin-bottom:20px;
}
```
